---
title: "Telecommunication Customer Churn Analysis"
output: word_document
---

## Authors 
Ali El-Sharif, Neha Panchal, Sagnik Adusumilli, Sarmad Shubber

## Background
 Nowadays, the telecom industry faces fierce competition in satisfying its customers. When a business loses customers, its profit and growth rate may be negatively affected. The churn rate of a company is percentage of subscribers to a service who discontinue their subscriptions within a given time period. 

## Objective
We will create a system that groups customers by their traits to better serve their individual needs.

## Data Analysis
The data has been sourced from Kaggle https://www.kaggle.com/jpacse/datasets-for-churn-telecom
The dataset is preprocessed and a balanced consisting of 71,047 instances and 58 attributes.

## Data Dictionary
Column Name               | Column Description  
--------------------------| ------------------- 
CustomerID	              | Unique customer ID
Churn	                    | Customer cut ties with service during a given time period (Yes/No)
MonthlyRevenue	          | Monthly revenue
MonthlyMinutes	          | Mean monthly minute of use
TotalRecurringCharge      | Mean total recurring charge
DirectorAssistedCalls     | Mean number of director assisted calls
OverageMinutes	          | Mean overage minute of use
RoamingCalls	            | Number of call were made and received outside of home area network operator
PercChangeMinutes	        | % change in minute of use
PercChangeRevenues        | % change in revenues
DroppedCalls	            | Number of telephone calls which, due to technical reasons were cut off
BlockedCalls	            | Number of incoming subscriber's blocked calls from specific telephone numbers
UnansweredCalls	          | Number of calls which are not picked up by an agent
CustomerCareCalls	        | Number of calls made by customer care to the subscriber.
ThreewayCalls	            | Number of conference call with two other two callers with local numbers
ReceivedCalls	            | Number of incoming calls that were answered
OutboundCalls	            | Number of  call initiated by a call center agent to a customer
InboundCalls	            | Number of call initiated by a  customer to call center agent
PeakCallsInOut	          | Mean of peak in and out voice calls
OffPeakCallsInOut	        | Mean number of in and out off peak voice calls
DroppedBlockedCalls	      | Mean number of dropped blocked calls
CallForwardingCalls	      | Number of redirected calls to another available telephone numbers
CallWaitingCalls	        | Number of suspended calls already in progress to accept a second call
MonthsInService	          | Months in Service
UniqueSubs	              | Number of unique subs
ActiveSubs	              | Number of active subs
ServiceArea	              | Area covered by the signal of a broadcasting station
Handsets	                | Outer part of mobile phone that does not include the battery or the sim card
HandsetModels	            | Models of mobile phones
CurrentEquipmentDays      | Number of days of the current equipment
AgeHH1	                  | Age of first HH member
AgeHH2	                  | Age of second HH member
ChildrenInHH	            | Presence of children in HH (Yes/No)
HandsetRefurbished	      | Pre-owned handset sent back due to a fault and has been repaired for re-sale (Yes/No)
HandsetWebCapable	        | Handset is web capable (Yes/No)
TruckOwner	              | Subscriber owns a Truck (Yes/No)
RVOwner	                  | Subscriber owns a recreational vehicle (Yes/No)
Homeownership	            | Subscriber owns a home (Yes/No)
BuysViaMailOrder	        | Buys via mail order (Yes/No)
RespondsToMailOffers  	  | Number of time Subscriber responds to a mail offers (Yes/No)
OptOutMailings	          | Chosen by not to be solicited by Mails (Yes/No)
NonUSTravel	              | Rravelled to non-US country (Yes/No)
OwnsComputer	            | Owns a personal computer (Yes/No)
HasCreditCard	            | Has a credit card (Yes/No)
RetentionCalls	          | Customer  made call to retention team
RetentionOffersAccepted   | Number of previous retention offers accepted
NewCellphoneUser	        | Known to be a new cell phone user	(Yes/No)
NotNewCellphoneUser	      | Known not to be a new cell phone user (Yes/No)
ReferralsMadeBySubscriber	| Number of referals made by subscriber
IncomeGroup	              | Subscriber income group
OwnsMotorcycle	          | Subscriber owns a	motor cycle (Yes/No)
AdjustmentsToCreditRating	| Number of adjustments were made to customer's credit rating (Up or Down)
HandsetPrice	            | Price of outer part of mobile phone not including  battery or sim card
MadeCallToRetentionTeam	  | Number of calls were made to retention team	(Yes/No)
CreditRating	            | Credit Rating
PrizmCode	                | PRIZM (Potential Rating Index for Zip Market) - census data market segmentation
Occupation	              | Occupation
MaritalStatus             | Marital status

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#import packages;
library(dplyr)
library(reshape2)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(mice)
library(VIM)
library(pROC)
library(caret)
library(sqldf)

# Clean all variables that might be left by other scripts
rm(list=ls(all=TRUE))

```


```{r, include=FALSE}
# Read data
telecomData = read.csv("../data/cell2celltrain.csv", header = TRUE, na.strings =  c("NA","","#NA"), sep=",")

```
## Data Exploartion
Summary:
```{r, echo = FALSE, warning = FALSE, message = FALSE}
summary(telecomData)

```
Structure:
```{r, echo = FALSE, warning = FALSE, message = FALSE}
str(telecomData)

```

Creating a correlation matrix
```{r fig.width=10, fig.height=10, echo = FALSE, warning = FALSE, message = FALSE}
# we can only Create a correlation matrix with only numeric data, therefore we are going to use sapply to only get numeric data
# sapply is applying a fuction over list of vector
numericData <- telecomData[sapply(telecomData,is.numeric)]
matrix <- cor(numericData, use="pairwise.complete.obs")
corrplot(matrix, type = "lower", method = "circle", order="hclust", tl.srt = 45, tl.cex = 0.7)
```
The given correlation matrix shows the correlation coefficients between sets of variables. This allows us to see which pairs have the highest correlation. A correlation matrix shows the correlation coefficients for combinations of 36 variables.
Positive correlations are displayed in blue, and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients. In the bottom side of the correlogram, the legend color shows the correlation coefficients and the corresponding colors.
In the above plot,
1) correlations with p-value > 0.01 are considered as insignificant. In this case, the correlation coefficient values are left blank or crosses are added.
2) )A positive correlation exists, when one variable increases as the other variable increase or one variable decreases while the other decrease.
For example: If you look at MonthlyMinutes and Monthly revenue variables, they are positively correlated. If there is an increase in monthly minutes, then it directly results in increased Monthly revenue.
3) A negative correlation, the variables move in inverse, or opposite, directions. In other words, as one variable increases, the other variable decreases.
For example: Customer ID and MonthsInService are negatively correlated. When new customer joins in, then he has not months in service. So there in no Correlation between these two variables and that’s why it is showing in dark red color in matrix.
4) A value of zero indicates that there is no relationship between the two variables.
In our case, dropped calls and CallForwarding calls variables are observed and found to have a correlation of 0; this means that there is no correlation, or relationship, between these two variables.
5) A weak correlation means that as one variable increase or decreases, there is a lower likelihood of being affected the second variable.
For example: DirectorAssisted calls and CurrentEquipment days are loosely correted.If there is one increase in directorassisted call,it doesn’t much affect CurrentEquipment days.




## Missing Data
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#lets check missing values

sort(colSums(is.na(telecomData)), decreasing = T)

```
Columns with NA: 
MonthlyRevenue, MonthlyMinutes, TotalRecurringCharge, DirectorAssistedCalls, OverageMinutes, RoamingCalls, PercChangeMinutes, PercChangeRevenues, Handsets, HandsetModels, CurrentEquipmentDays, AgeHH1, AgeHH2.

The maximum number of records with NA columns is 909 for columns AgeHH1 and AgeHH2. We are going to eliminate these rows with NA values.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#We can skip all the data with missing values:
telecomData <- na.omit(telecomData)
```
After running the na.omit command, number of rows was 51,047 and now it is 49,752 rows.


```{r feature_distribution, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap=" Distribution of Important Features"}

tmp = telecomData %>%filter(complete.cases(.))

library(RColorBrewer) # color palettes
# pick palettes 
mainPalette = brewer.pal(8,"Dark2")

p1 = tmp %>% ggplot(aes(x=DirectorAssistedCalls  )) + geom_density(fill=mainPalette[1], colour=mainPalette[1], alpha = 0.2) + scale_x_log10()

p2 = tmp %>% ggplot(aes(x=RoamingCalls  )) + geom_density(fill=mainPalette[2], colour=mainPalette[2], alpha = 0.2) + scale_x_log10()

p3 = tmp %>% ggplot(aes(x=DroppedCalls  )) + geom_density(fill=mainPalette[3], colour=mainPalette[3], alpha = 0.2) + scale_x_log10()

p4 = tmp %>% ggplot(aes(x=BlockedCalls  )) + geom_density(fill=mainPalette[4], colour=mainPalette[4], alpha = 0.2) + scale_x_log10()

p5 = tmp %>% ggplot(aes(x=UnansweredCalls  )) + geom_density(fill=mainPalette[5], colour=mainPalette[5], alpha = 0.2) + scale_x_log10()

p6 = tmp %>% ggplot(aes(x=CustomerCareCalls  )) + geom_density(fill=mainPalette[6], colour=mainPalette[6], alpha = 0.2) + scale_x_log10()

p7 = tmp %>% ggplot(aes(x=ThreewayCalls  )) + geom_density(fill=mainPalette[7], colour=mainPalette[7], alpha = 0.2) + scale_x_log10()

p8 = tmp %>% ggplot(aes(x=ReceivedCalls  )) + geom_density(fill=mainPalette[8], colour=mainPalette[8], alpha = 0.2) + scale_x_log10()

p9 = tmp %>% ggplot(aes(x=OutboundCalls  )) + geom_density(fill=mainPalette[1], colour=mainPalette[1], alpha = 0.2) + scale_x_log10()

p10 = tmp %>% ggplot(aes(x=InboundCalls  )) + geom_density(fill=mainPalette[2], colour=mainPalette[2], alpha = 0.2) + scale_x_log10()

p11 = tmp %>% ggplot(aes(x=PeakCallsInOut  )) + geom_density(fill=mainPalette[3], colour=mainPalette[3], alpha = 0.2) + scale_x_log10()

p12 = tmp %>% ggplot(aes(x=OffPeakCallsInOut  )) + geom_density(fill=mainPalette[4], colour=mainPalette[4], alpha = 0.2) + scale_x_log10()

p13 = tmp %>% ggplot(aes(x=DroppedBlockedCalls  )) + geom_density(fill=mainPalette[5], colour=mainPalette[5], alpha = 0.2) + scale_x_log10()

p14 = tmp %>% ggplot(aes(x=CallForwardingCalls  )) + geom_density(fill=mainPalette[6], colour=mainPalette[6], alpha = 0.2) + scale_x_log10()

p15 = tmp %>% ggplot(aes(x=CallWaitingCalls  )) + geom_density(fill=mainPalette[7], colour=mainPalette[7], alpha = 0.2) + scale_x_log10()

p16 = tmp %>% ggplot(aes(x=RetentionCalls  )) + geom_density(fill=mainPalette[8], colour=mainPalette[8], alpha = 0.2) + scale_x_log10()


library(gridExtra) # arrange grids

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8)
grid.arrange(p9,p10,p11,p12,p13,p14,p15,p16)

rm(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13,p14,p15,p16,tmp)
```



## Insights from Data explorations

## Data Preparation

# create normalization function and apply to the data
normalize <-function(x){
  return ((x-min(x))/(max(x)-min(x)))
}
telecomData_n <- as.data.frame(lapply(telecomData,normalize))

# apply R's z-score standarization function to the data
telecomData_z <-as.data.frame(lapply(telecomData,scale))

# create and apply threshold function
threshold <- function(x,c) {
  return (ifelse(x > c,1,0))
}
threshold_b <- as.data.frame(lapply(telecomData, threshold, c=1))


## Modeling

```{r, echo = FALSE, warning = FALSE, message = FALSE}
telecomDataSmall <- telecomData[1:10000,]

#Hierarchical Clustering
#We need to calculate how far away each customer is from the cluster’s mean. To do this we could use 
#many distances/dissimilarity index, one of which is the Gower dissimilarity. We use Gower because we have a mix of numerical and categorical data

library(cluster)
D=daisy(telecomDataSmall, metric='gower')

#After the creation of a distance matrix, we implement a Ward’s hierarchical clustering procedure:
  
H.fit <- hclust(D, method="ward.D2")
plot(H.fit) # display dendrogram

groups <- cutree(H.fit, k=4) # cut tree into 4 clusters

# draw dendogram with red borders around the 4 clusters
rect.hclust(H.fit, k=4, border="red") 


# 2D representation of the Segmentation:
clusplot(telecomDataSmall, groups, color=TRUE, shade=TRUE,
         labels=2, lines=0, main= 'Customer segments')

telecomDataSmallWithCluster <- telecomDataSmall
telecomDataSmallWithCluster$Cluster<-groups

# telecomDataSmallWithCluster is our dataset with a column Cluster that indicates which cluster each customer belongs to, this can be used for further analysis


#Convert factor columns to numeric in order to do the clustering.
telecomDataSmallNumeric <- telecomDataSmall
# Convert Factor columns to numeric columns
for (i in colnames(telecomDataSmallNumeric)){
  if (is.factor(telecomDataSmallNumeric[,c(i)])){
    telecomDataSmallNumeric[,c(i)] <- as.numeric(telecomDataSmallNumeric[,c(i)])
  }
}

remove(i)

#scale the variables
scaled_telecomDataSmallNumeric <- scale(telecomDataSmallNumeric)

D=daisy(scaled_telecomDataSmallNumeric, metric='gower')

#After the creation of a distance matrix, we implement a Ward’s hierarchical clustering procedure:
  
H.fit <- hclust(D, method="ward.D2")

groups <- cutree(H.fit, k=4) # cut tree into 4 clusters

#pca
pcmp <- princomp(telecomDataSmallNumeric)
pred_pc <- predict(pcmp, newdata=scaled_telecomDataSmallNumeric)[,1:2]
#pred_pc
comp_dt <- cbind(as.data.table(pred_pc),cluster = as.factor(groups), Labels = telecomDataSmall$CustomerID)
#comp_dt
ggplot(comp_dt,aes(Comp.1,Comp.2))+
  geom_point(aes(color = cluster),size=3) #+xlab('x')+ylab('density')


#kmeans
kclust <- kmeans(scaled_telecomDataSmallNumeric,centers = 4,iter.max = 100)
#kclust
ggplot(comp_dt,aes(Comp.1,Comp.2))+
  geom_point(aes(color = as.factor(kclust$cluster)),size=3)

library(fpc)
tunek <- kmeansruns(scaled_telecomDataSmallNumeric,krange = 1:5,criterion = "ch")
tunek$bestk #2

tunekw <- kmeansruns(scaled_telecomDataSmallNumeric,krange = 1:3,criterion = "asw")
tunekw$bestk #2

```