---
title: "main"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#import packages;
library(dplyr)
library(reshape2)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(mice)
library(VIM)
library(pROC)
library(caret)
library(sqldf)
```

## Abstract

TBD

## Background 
TBD



## Objective
TBD

# Data Analysis
Source of data: [Moro et al., 2014] S. Moro, P. Cortez and P. Rita. A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31, June 2014

Extracted from https://archive.ics.uci.edu/ml/datasets/Bank+Marketing#



## Data Dictionary
Datat set is comprised of database of contributions made at both the federal level and in 13 provinces and territories.

Column Name            | Column Description  
-----------------------| ------------------- 
age                    | Age of the customer
job                    | Customer's Occupation*
marital                | Marital status of the customer* 
education              | Education level of the customer*
default                | Has credit in default (yes/no)
balance                | ??
housing                | has housing loan (yes/no)
loan                   | has personal loan (yes/no)  
contract               | contact communication type (categorical: 'cellular','telephone') 
day                    | ??
month                  | month the customer was contrained previously in the year
duration               |  last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then                             y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously 
campaign               |  number of contacts performed during this campaign and for this client (numeric, includes last contact)
pdays                  | number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously                           contacted)
previous               | number of contacts performed before this campaign and for this client
poutcome               | outcome of the previous marketing campaign
y                      | has the client subscribed a term deposit? (binary: 'yes','no')

## Data Exploartion
```{r}
bankData = read.csv("../Data/bank.csv", header = TRUE, na.strings =  c("NA","","#NA"), sep=";")
```


## Missing Data

## Insights from Data explorations

## Data Preparation

## Data Imputing

## Feature Selection

## Modeling

You can also embed plots, for example:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
