---
title: "main"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#import packages;
library(dplyr)
library(reshape2)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(mice)
library(VIM)
library(pROC)
library(caret)
library(sqldf)
```

## Abstract
Nowadays, telecom industry faces fierce competition in satisfying its customers. When a business loses customers, its profit and growth rate may be negatively affected. The churn rate of a company is percentage od subscribers to a service who discontinue thier subscritions within a given time period. We will try to create a sytem will group customers by thier traits so indiviudal nee
## Background 
TBD



## Objective
TBD

# Data Analysis
Source of data: https://www.kaggle.com/jpacse/datasets-for-churn-telecom



## Data Dictionary
Datat set is comprised of database of contributions made at both the federal level and in 13 provinces and territories.

Column Name            | Column Description  
-----------------------| ------------------- 


## Data Exploartion
```{r}
telecomData = read.csv("../data/cell2celltrain.csv", header = TRUE, na.strings =  c("NA","","#NA"), sep=",")
summary(telecomData)
```

Creating a correlation matrix
```{r fig.width=10, fig.height=10}
library(corrplot)
library(functional)
# we can only Create a correlation matrix with only numeric data, therefore we are going to use sapply to only get numeric data
# sapply is applying a fuction over list of vector
numericData <- telecomData[sapply(telecomData,is.numeric)]
matrix <- cor(numericData, use="pairwise.complete.obs")
corrplot(matrix, type = "lower", method = "circle", order="hclust", tl.srt = 45, tl.cex = 0.7)
```
## Missing Data

## Insights from Data explorations

## Data Preparation

## Data Imputing

## Feature Selection

## Modeling

You can also embed plots, for example:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
